---
title: "Sex_comparison_general_linear_model"
author: "Tyler Audet"
date: "2024-04-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script is an example of the linear modelling for the male vs. female modelling. This was run for each treatment independantly on a remote computer.

First we read in a frequency table of major and minor allele frequencies. These frequency tables are not provided due to size, but are generated using the frequency table functions in 'grenedalf'.

```{r}
library(data.table)
library(tidyverse)
library(emmeans)


majAllele <- fread("../data/C1.csv")

header <- c("chrom", "pos", "maj_F", "min_F", "maj_M", "min_M")

colnames(majAllele) <- header

test <- as.data.frame(pivot_longer(majAllele,
                                   cols = maj_F:min_M,
                                   names_to = "treatment"))

test2 <- separate(test, col = treatment, sep = "_", into = c("allele","sex"))
test3 <- as.data.frame(pivot_wider(test2, names_from = allele, values_from = value))

test3$pos <- as.factor(test3$pos)


```

Next we iterate over all positions with a binomial glm with a logit link function. We extract the emmeans and the confidence intervals and output them into their own table.

```{r}
out <- matrix(nrow = length(unique(test3$pos)), ncol = 9)


x <- 1

count <- 1
for(i in unique(test3$pos)) {
  
  count <- count + 1
  if (count%%10 == 0){
    print(count)
    print(Sys.time())
  }
  
  pos <- i
  data <- test3[(test3$pos == i),]
  mod1 <- glm(cbind(maj,min) ~ sex,
              data = data, family = binomial(link = "logit"))
  
  mod_emm <- emmeans(mod1, ~sex, type = "response")
  #mod_emm_df <- as.data.frame(as.data.frame(mod_emm))

contrast <- pairs(mod_emm)
contrast_conf <- confint(contrast)
contrast_df <- as.data.frame(as.data.frame(contrast))
contrast_df_conf <- as.data.frame(as.data.frame(contrast_conf))

num_contrast <- contrast_df[,2:7]
num_conf <- contrast_df_conf[,5:6]

contrast_final <- as.matrix(cbind(i,num_contrast,num_conf))


out[x,] <- contrast_final

x <- x+1

}

colnames(out) <- c("pos","odds.ratio","SE","df","null","z.ratio","pval","UCL","LCL")

fwrite(out, "./C1_sex_model.txt", quote = FALSE, row.names = FALSE, col.names = TRUE, sep = "\t")

```

